PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-01-01 txt/maths.tex
prelinria done

\subsubsection{Mathematical preliminaries}

%
%
% The action of
%quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
%seen as  a measure-and-prepare quantum channel as follows
%$
%\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
%$

\subsubsection{Diamond norm and the task of discrimination between quantum operators}
This section defines a special norm on the space of linear mappings, known as
the diamond norm, and establishes some of its properties.
Next, we will explain the precise connection between the diamond
norm and the task of quantum operation discrimination.

We define the diamond norm of a map $\Phi$ as
\begin{equation}
\|\Phi\|_\diamond = \max_{\| \ket{\psi}\|_1=1} \| \left(\Phi \otimes \1\right) (\proj{\psi}) \|_1.
\end{equation}
The quantum state $\ket{\psi}$ which maximizes the diamond norm we will call as discriminator.

The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum measurements, $\PP$  and $\mathcal{Q}$,
in terms of their distance with the use of the diamond norm
\begin{equation}
p_{\text{succ}}(\PP, \mathcal{Q}) =  \frac12 + \frac14 \| \PP - \mathcal{Q} \|_\diamond.
\end{equation}

\subsubsection{Example of computing optimal measurements and discriminator}

So far, we only discussed how the discrimination is performed using two different
schemes, assuming, that all needed components ($V_0$, $V_1$,$\ket{\psi_0}$) are known.
In this section, we will show how all the necessary ingredients look like for
von Neumann measurements defined by the parametrized Fourier family:

\begin{equation}
	U = H
	\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger
\end{equation}

where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.

@@

\subsubsection{Conditional binary measurement/Optimal measurement}\label{sec_example_final_measurement}

\begin{proposition}
The  optimal measurement has the following form
\begin{equation}
\begin{split}
\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1
\proj{0} V_1^\dagger  \\
\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1
\proj{1} V_1^\dagger,
\end{split}
\end{equation}
where for each $\phi \in \mathbb{R}$ and  the unitaries $V_0$,  $V_1$
have the following form
\begin{equation}
V_0 = \left(\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i
\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi -
\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)\end{array}\right),
\end{equation}
\begin{equation}
V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i
\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} \right)
&  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right).
\end{equation}
\end{proposition}

>>>>

\subsubsection{Optimal discrimination scheme}

The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum measurements, $\PP$  and $\mathcal{Q}$,
in terms of their distance with the use of the diamond norm
\begin{equation}
p_{\text{succ}}(\PP, \mathcal{Q}) =  \frac12 + \frac14 \| \PP - \mathcal{Q} \|_\diamond, 
\end{equation}
where 
\begin{equation}
\|\PP\|_\diamond = \max_{\| \ket{\psi}\|_1=1} \| \left(\PP \otimes \1\right) (\proj{\psi}) \|_1.
\end{equation}
The quantum state $\ket{\psi}$ which maximizes the diamond norm is called as discriminator.

Furthermore,  thanks to the proof of the Holevo-Helstrom theorem, it is possible to construct the necessary components  ($\ket{\psi_0}$, $V_0$, $V_1$) to create the optimal discrimination strategy.

\subsubsection{Discrimination scheme for parameterized Fourier family}

So far, we only discussed how the discrimination is performed using two different
schemes, assuming that all needed components ($\ket{\psi_0}$, $V_0$, $V_1$) are known. While the determination of $V_0$ and $V_1$ is a simple task provided the discriminator $\ket{\psi_{0}}$ is known, it turns out, 
sad to say, determining the analytical form of the discriminator $\ket{\psi_{0}}$
is, in general, NP-hard problem. 
We have found the analytical solution of the discrimination scheme for a particular class of von Neumann measurements.

Here,  we will show how all the necessary ingredients look like for
von Neumann measurements defined by the parameterized Fourier family:

\begin{equation}
	U = H
	\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger, 
\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.

@@

\begin{proposition}
The  optimal measurement has the following form
\begin{equation}
\begin{split}
\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1
\proj{0} V_1^\dagger,   \\
\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1
\proj{1} V_1^\dagger,
\end{split}
\end{equation}
where for each $\phi \in \mathbb{R}$ and  the unitaries $V_0$,  $V_1$
have the following form
\begin{equation}
V_0 = \left(\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i
\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi -
\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)\end{array}\right),
\end{equation}
\begin{equation}
V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i
\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} \right)
&  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right).
\end{equation}
\end{proposition}
